---
title: "WUP_04"
author: "Audrea Pena"
date: "fall 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
mydata<-read.csv(file="GSScsv.csv")
```
```{r}
boxplot(WORDSUM~DEGREE,data=mydata)
```  

The middle box and the far right box in the plot show that they are below 50% in comparison to the graduate degree. The means of the two sets are different from each other, therefore if you run any test with the two boxes (the one in the middle and the one on the far right), you can suspect that mean values for the WORDSUM test are a lower value than the boxes labeled "Bachelor", "Graduate" and "Junior College".  

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```
  The individuals that have a junior college, bachelors degree and a graduate degree have a higher mean result than those with a high school and middle school category.  According to the plot, the individuals with a degree have a mean value of 9 or higher. While the high school and junior high school students have a mean value of 8 or less.
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```  

  The P value of 0.000236 is extremely small. The P value represents the probability of getting this result or a farther with NULL hypothesis being actually true. i do not believe this to be true, therefore I reject the NULL hypothesis.
  
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1, conf.level=.95), las=2)
```

  This shows that high school and graduate are different from each other, and that LT high school and graduate are different as well. We can see that LT high school-graduate and high school-graduate do no t hit the "0" line.
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

  In this boxplot we discussed how individuals with a higher degree of education, score  higher WORDSUM scores than those individuals in the LT High School and High School category. We established that individuals that had a college degree had a mean value of 9 or more, while the LT High School and High School had a mean value of 8 or less. We found that our P value was so small 0.00236, that the NULL hypothesis was not true, therefore we rejected the NULL hypothesis. Using the TukeyHSD model we were able to elaborate further at a 95% confidence level LT High School-graduate and High School-graduate categories did not reach the "0" mark. In conclusion, these two categories were different. 